---
title: "Youth Risk Behaviour Analysis"
author: "Matias Pere"
date: "15/10/2020"
output: html_document

---


```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, setup2, warning=FALSE, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)
# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, echo=FALSE, include=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(GGally)
library(readxl)
library(here)
library(skimr)
library(janitor)
library(broom)
library(tidyquant)
library(infer)
library(openintro)
library(tidyquant)
```



Every two years, the Centers for Disease Control and Prevention conduct the [Youth Risk Behavior Surveillance System (YRBSS)](https://www.cdc.gov/healthyyouth/data/yrbs/index.htm) survey, where it takes data from high schoolers (9th through 12th grade), to analyze health patterns. You will work with a selected group of variables from a random sample of observations during one of the years the YRBSS was conducted.

## Loading the data

```{r}
data(yrbss)
glimpse(yrbss)
skim(yrbss)
```

## Exploratory Data Analysis

We will first start with analyzing the `weight` of participants in kilograms.

There are 1004 missing values in weight, and distribution of weight is right skewed.

```{r, eda_on_weight}
yrbss %>% 
  filter(is.na(weight)) %>% 
  summarize(missing_count = n()) #number of missing observations in weight

yrbss %>% 
  ggplot(aes(weight)) +
  geom_histogram() + #plot the distribution of weight
  theme_bw() + #theme
  labs(title = "distribution of weight") # add title
```

Next, we consider the possible relationship between a high schooler’s weight and their physical activity.

Let’s create a new variable `physical_3plus`, which will be `yes` if they are physically active for at least 3 days a week, and `no` otherwise.

  
```{r}
yrbss <- yrbss %>% 
  mutate(physical_3plus = ifelse(physically_active_7d >= 3, "yes", "no"))

yrbss %>% filter(!is.na(physical_3plus)) %>% 
  group_by(physical_3plus) %>% 
  summarise(count = n()) %>% 
  mutate(prop= count/sum(count),
         se = sqrt(prop*(1-prop)/count),  #calculate stand error
         t_critical = qt(0.975, count-1), #calculate t critical for 95% CI
         CI_low = prop - t_critical*se,
         CI_high = prop + t_critical*se)  #calculate high end for 95% CI)

```

The  95% confidence interval for the population proportion of high schools that are *NOT* active 3 or more days per week is 0.317 to 0.345.
  

In the following we make a boxplot of `physical_3plus` vs. `weight`. 

```{r, boxplot}
yrbss %>% 
  filter(!is.na(physical_3plus)) %>% 
  ggplot(aes(x = physical_3plus, y = weight, color = physical_3plus)) +
  geom_boxplot(show.legend = FALSE) +
  labs(title = "boxplot of physical_3plus vs. weight") +
  theme_bw()
```
The median `weight` of `physical_3plus==yes` is larger, and there are more extreme `weight` of `physical_3plus==no`. I expected the median `weight` of `physical_3plus==yes` to be smaller because they tended to consume more energy, however that is not what the data shows. Maybe this is because people who do sports 3 times a week or more, are muscular and therefore heavy. Muscles are heavier than fat.

## Confidence Interval

Boxplots show how the medians of the two distributions compare, but we can also compare the means of the distributions using either a confidence interval or a hypothesis test.

```{r}
yrbss %>%
  group_by(physical_3plus) %>%
  filter(!is.na(physical_3plus)) %>% 
  summarise(mean_weight = mean(weight, na.rm = TRUE),
            sd_weight = sd(weight, na.rm=TRUE),
            count = n(),
            se_weight = sd_weight/sqrt(count),
            t_critical = qt(0.975, count-1), 
            margin_of_error = t_critical * se_weight,
            lower = mean_weight - t_critical * se_weight,
            upper = mean_weight + t_critical * se_weight
            )

```

There is an observed difference of about 1.77kg (68.44 - 66.67), and we notice that the two confidence intervals do not overlap. It seems that the difference is at least 95% statistically significant. Let us also conduct a hypothesis test.

## Hypothesis test with formula

Null Hypothesis (h0): There is no difference of mean weights for students who exercise compared to those who do not.
Alternative Hypothesis (h1): There is a difference of mean weights.

```{r}
t.test(weight ~ physical_3plus, data = yrbss)
```



## Hypothesis test with `infer`

Next, we will introduce a new function, `hypothesize`, that falls into the infer workflow. We will use this method for conducting hypothesis tests.

But first, we need to initialize the test, which we will save as `obs_diff`.

```{r}
obs_diff <- yrbss %>%
  specify(weight ~ physical_3plus) %>%
  calculate(stat = "diff in means", order = c("yes", "no"))

```

After we have initialized the test, we need to simulate the test on the null distribution, which we will save as null.


```{r}
null_dist <- yrbss %>%
  specify(weight ~ physical_3plus) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("yes", "no"))

```

We can visualize this null distribution with the following code:

```{r}
ggplot(data = null_dist, aes(x = stat)) +
  geom_histogram()

```


Now that the test is initialized and the null distribution formed, we can visualise to see how many of these null permutations have a difference of at least `obs_stat` of `r obs_diff %>% pull() %>% round(2)`?

We can also calculate the p-value for your hypothesis test using the function `infer::get_p_value()`.

```{r}

null_dist %>% visualize() +
  shade_p_value(obs_stat = obs_diff, direction = "two-sided")

null_dist %>%
  get_p_value(obs_stat = obs_diff, direction = "two_sided")

```

This is the standard workflow for performing hypothesis tests.
As the p-value is 0 we can conclude to reject our null-hypothesis.